---
title: "Computing Fourier Series Numerically For Simple Functions Using R"
author: "Lewis Peaty"
date: "25 March 2016"
output: html_document
---

Here is an example of how to compute and plot Fourier Series for simple functions using only the standard R library.

```{r}

# try using some of these functions instead
# f<- function(x){cos(x)}
# f<- function(x){x>-0.5 & x<0.5}
# f<- function(x){x>0}

# edit these values to change the function, interval, or the number of harmonics, 
f <- function(x){x^3 + 0.5 * x^2 + 0.2 * x + 0.1} # the function we want to find the fourier series of
lower <- -1 # lower bound
upper <- 1 # upper bound
increment <- 0.01 # time increment (smaller -> slower)
num_harmonics <- 400 # number of harmonics to use- try a small value for fun!

t <- seq(from=lower,to=upper,by=increment) # vector of time values to evaluate f over
f_t <- sapply(t,f) # evaluate f over t
f0 <- 1 / (upper - lower)   # fundamental frequency of f
a_n <- as.numeric(0) # vector of cosine coefficients
b_n <- as.numeric(0) # vector of sine coefficients
a_n[1] <- sum(f_t) / 2 # compute a_0 by integrating f_t / 2. This looks a little odd because R does not zero-index vectors
for(n in 1:num_harmonics){ # compute coefficients for each harmonic in turn
  w <- f0 # shouldn't there be 2*pi here??
  # compute Cosine Fourier Coefficient a_n by integrating f * cos(n w x)
  c_n <- sapply(t,function(x){cos(n*w*x)})
  a_n <- append(a_n, sum(f_t * c_n))
  # compute Sine Fourier Coefficient b_n by integrating f * sin(n w x)
  s_n <- sapply(t,function(x){sin(n*w*x)})
  b_n <- append(b_n, sum(f_t * s_n))
}

```

The coefficients have been computed. They can be tabulated and displayed (only showing the first 20 here).
Note that everything is offset by 1 because R starts vector indices at 1 not 0. E.g. a\_n[1] is actually coefficent a_0:

```{r}

coefs <- cbind(a_n,b_n)
if(length(coefs[,1])>20){print(coefs[1:20,])}else{print(coefs)}

```

The Fourier Series can be plotted in the time domain to see how it compares with the original function:

```{r}

# Time to plot the harmonics!
F <- function(coefs,f0,ts){
  w <- f0 # shouldn't there be 2*pi here??
  n_coefs <- nrow(coefs)
  accumulator <- numeric(length(ts))
  accumulator <- accumulator + coefs[1,1] # a_0
  for(n in 1:n_coefs-1){ # R vectors are not zero-indexed so this part looks a bit funny
    accumulator <- accumulator + sapply(ts,function(x){coefs[n+1,1]*cos(n*w*x)+coefs[n+1,2]*sin(n*w*x)})
  }
  accumulator
}

```

Plotting the original function:

```{r}

plot(t,f_t,"l")

```

Plotting the computed Fourier Series:

```{r}

plot(t,F(coefs,f0,t),"l")


```

