---
title: "T Test Presentation"
author: "Lewis Peaty"
date: "13 May 2016"
output: ioslides_presentation
---

#Let's learn about T-Tests!

## What Is The T-Test?

> - The T-Test detects FALSE POSITIVES.
> - The T-Test can tell you when your ideas (hypotheses) are wrong.

## For Example:


> * We think that "300M_FROM_SALT_LAKE" implies an increase in corrosion rate for asset X.    
> * Follow the procedure:   
  > * Randomly select equal sized samples of asset X which are <=300M from a salt lake, and >300M from a salt lake.
  > * Run the test.
  > * If the T-Test passes then you may say there is evidence "300M_FROM_SALT_LAKE" really does imply an increase in corrosion rate at the given significance level!

## An Example Population Distribution

Let's start by looking at a population of assets.

```{r,echo=FALSE}
`%c%` <- function(a,b){paste(sep="",a,b)}
set.seed(23423544)
dist <- (round(x = rnorm(1000,mean = 50,sd = 3),digits = 0))
dist <- c(dist,round(x = rnorm(1000,mean = 75,sd = 3)))
dist <- c(dist,round(x = runif(n = 1000,min = 0,max = 100)))
dist <- ifelse(dist<0,0,dist)
`Population Distribution` <- dist
hist(`Population Distribution`,ylab = "Count of Asset X", xlab = "Corrosion Rate, n = 3000, mean = 58.305, sd = 20.72153")
# print("The population mean is:")
# print("n = " %c% length(dist))
# print(mean(dist))
# print(sd(dist))

```



## Sample Near To Salt Lake



```{r, echo=FALSE}
dist_df <- as.data.frame(dist)
dist_df$`300M_FROM_SALT_LAKE` <- sample(c(TRUE,FALSE),size = 3000,replace = TRUE)
dist_df$`300M_FROM_COAST` <-ifelse(dist_df$dist>50,FALSE,TRUE)
for(i in 1:nrow(dist_df)){
  if(dist_df[i,1]>50)
    {dist_df$`300M_FROM_COAST`[i] <- sample(c(TRUE,FALSE),size = 1,replace=TRUE,prob=c(.75,.25))}
  else
    {dist_df$`300M_FROM_COAST`[i] <- sample(c(TRUE,FALSE),size = 1,replace=TRUE,prob=c(.25,.75))}
}
s1 <- sample(dist_df$dist[dist_df$`300M_FROM_SALT_LAKE`==TRUE],size = 50,replace = FALSE)
s2 <- sample(dist_df$dist[dist_df$`300M_FROM_SALT_LAKE`==FALSE],size = 50,replace = FALSE)

s3 <- sample(dist_df$dist[dist_df$`300M_FROM_COAST`==TRUE],size = 100,replace = FALSE)
s4 <- sample(dist_df$dist[dist_df$`300M_FROM_COAST`==FALSE],size = 100,replace = FALSE)
`Sample Distribution 1` <- s1
`Sample Distribution 2` <- s2
`Sample Near Salt Lake` <- s1
`Sample Far From Salt Lake` <- s2
`Sample Near Coast` <- s3
`Sample Far From Coast` <- s4
# mean(s1)
# sd(s1)
# mean(s2)
# sd(s2)
hist(`Sample Near Salt Lake`,ylab = "Count of Asset X", xlab = "Corrosion Rate, n = " %c% length(s1) %c% ", mean = " %c% mean(s1) %c% ", sd = " %c% sd(s1) )
```

## Sample Far From Salt Lake

```{r,echo=FALSE}

hist(`Sample Far From Salt Lake`,ylab = "Count of Asset X", xlab = "Corrosion Rate, n = " %c% length(s2) %c% ", mean = " %c% mean(s2) %c% ", sd = " %c% sd(s2) )


```

## Perform The T-Test

Invoke the function in R (or Excel!), feeding in the two sample populations 

```{r}

t.test(s1,s2)

```

## What does it mean?

> - P-Value = 0.6214 (62.14%)
> - There is a 62.14% chance the variation in corrosion rates between our two samples was due to random chance!
> - I.e. Our hypothesis that being <300M from a salt lake increases corrosion rate is probably wrong!


## Sample Near To Coast

Salt lakes are obviously irrelevant.
Let's try assets which are within "300M_TO_COAST" instead!


```{r,echo=FALSE}

hist(`Sample Near Coast`,ylab = "Count of Asset X", xlab = "Corrosion Rate, n = " %c% length(s3) %c% ", mean = " %c% mean(s3) %c% ", sd = " %c% sd(s3) )


```

## Sample Far From Coast

The histograms look different... But let's quantify the difference with a T-Test.

```{r,echo=FALSE}

hist(`Sample Far From Coast`,ylab = "Count of Asset X", xlab = "Corrosion Rate, n = " %c% length(s4) %c% ", mean = " %c% mean(s4) %c% ", sd = " %c% sd(s4) )


```

## T-Test For 300M From Coast

Invoke the function again

```{r}

t.test(s3,s4)

```

## Results

> - Wow! A p-value of 0.0001098 means there is a 0.01% chance that the difference in corrosion rates is due to random chance.

> - I.e We can now claim we have STRONG EVIDENCE that assets within 300M of the coast corrode faster than other assets of the same type. 

## Final Note

BEWARE: P-Values can be manipulated by using small sample sizes.

